---
title: "database fisher test"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
name: "Adrian Cantu"
header-includes:
    - \usepackage[labelformat=empty]{caption}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library("tidyverse"))
library(kableExtra)
source("fisher_exact_test.R")
library(pander)
#driver_level_min=1
intSites <- readRDS("data/intSites_plus0.rds")
intSitesTCGA_1 <- data.frame(readRDS('data/intSitesTCGA_1.rds'))
intSitesTCGA_2 <- data.frame(readRDS('data/intSitesTCGA_2.rds'))
intSitesTCGA_3 <- data.frame(readRDS('data/intSitesTCGA_3.rds'))
intSitesTCGA_4 <- data.frame(readRDS('data/intSitesTCGA_4.rds'))
```


Fisher exact p-test was performed to compare the prevalence of insertion sites near oncogenes in pre-infusion samples (time 0) versus samples at time point ~6 years. We consider two ways of counting the insertion sites, Simple counts (where each observed insertion site counts only once) and Sonic abundance counts (where each observed insertion site counts a number of times equal to its sonic abundance).


```{r sample_table, echo=FALSE}
sample_table <- read.csv('summary_all.csv')
used <- c('GTSP0886','GTSP0890','GTSP4185','GTSP3310','GTSP3311','GTSP3438')
sample_table %>% 
  filter(GTSP %in% used) %>%
  transmute(GTSP=GTSP,SampleName=c('P1_t72','P2_t72','P3_t72','P5_t0_1','P5_t0_2','P5_t70'),Timepoint=Timepoint,patient=sub('^p', '',patient),InferredCells=InferredCells,UniqueSites=UniqueSites ) %>%
kable( "latex", booktabs = TRUE, longtable = TRUE, caption = "Used samples") %>%
  kable_styling(latex_options = c("hold_position"))


```

For oncogenes definitions (Databases) we use the "Bushman lab oncogenes database" and four levels of The Cancer Genome Atlas (TCGA) version of the OncoVar database. Each gene in the OncoVar database is assigned a "driver level" score, a number 0-4 that quantifies the evidence for a gene being oncogenic (a level 4 gene is most likely oncogenic, while a level 1 only has little evidence and a level 0 none). Significant values for the Fisher test are marked red.


\newpage
```{r b_lab_data, echo=FALSE, results='asis'}

current_intsites <- intSites
pandoc.header("Busman lab database", level = 1)

p_val_table <- get_pval_table(current_intsites)
cont_table <- get_cont_table(current_intsites)

p_val_table_w <- get_pval_table_w(current_intsites)
cont_table_w <- get_cont_table_w(current_intsites)


t_p_val_table <- p_val_table %>% kable( booktabs = TRUE, digits=5, caption = "pvalue simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table$P5_t0_1<=0.05,'red','black'))
t_p_val_table_w <- p_val_table_w %>% kable( booktabs = TRUE, digits=5, caption = "pvalue Sonic abundance") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table_w$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table_w$P5_t0_1<=0.05,'red','black'))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t_p_val_table,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t_p_val_table_w,
    "\\end{minipage} 
\\end{table}"
))  


cont_table %>% kable( booktabs = TRUE, digits=3, caption = "simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "simple counts proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))

cont_table_w %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table_w %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cat('\\newpage')
```


```{r oncovar_1, echo=FALSE, results='asis'}
current_intsites <- intSitesTCGA_1
pandoc.header("OncoVar lvl 1", level = 1)

p_val_table <- get_pval_table(current_intsites)
cont_table <- get_cont_table(current_intsites)

p_val_table_w <- get_pval_table_w(current_intsites)
cont_table_w <- get_cont_table_w(current_intsites)


t_p_val_table <- p_val_table %>% kable( booktabs = TRUE, digits=5, caption = "pvalue simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table$P5_t0_1<=0.05,'red','black'))
t_p_val_table_w <- p_val_table_w %>% kable( booktabs = TRUE, digits=5, caption = "pvalue Sonic abundance") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table_w$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table_w$P5_t0_1<=0.05,'red','black'))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t_p_val_table,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t_p_val_table_w,
    "\\end{minipage} 
\\end{table}"
))  


cont_table %>% kable( booktabs = TRUE, digits=3, caption = "simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "simple counts proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))

cont_table_w %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table_w %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cat('\\newpage')
```


```{r oncovar_2, echo=FALSE, results='asis'}
current_intsites <- intSitesTCGA_2
pandoc.header("OncoVar lvl 2", level = 1)

p_val_table <- get_pval_table(current_intsites)
cont_table <- get_cont_table(current_intsites)

p_val_table_w <- get_pval_table_w(current_intsites)
cont_table_w <- get_cont_table_w(current_intsites)


t_p_val_table <- p_val_table %>% kable( booktabs = TRUE, digits=5, caption = "pvalue simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table$P5_t0_1<=0.05,'red','black'))
t_p_val_table_w <- p_val_table_w %>% kable( booktabs = TRUE, digits=5, caption = "pvalue Sonic abundance") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table_w$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table_w$P5_t0_1<=0.05,'red','black'))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t_p_val_table,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t_p_val_table_w,
    "\\end{minipage} 
\\end{table}"
))  


cont_table %>% kable( booktabs = TRUE, digits=3, caption = "simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "simple counts proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))

cont_table_w %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table_w %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cat('\\newpage')
```

```{r oncovar_3, echo=FALSE, results='asis'}
current_intsites <- intSitesTCGA_3
pandoc.header("OncoVar lvl 3", level = 1)

p_val_table <- get_pval_table(current_intsites)
cont_table <- get_cont_table(current_intsites)

p_val_table_w <- get_pval_table_w(current_intsites)
cont_table_w <- get_cont_table_w(current_intsites)


t_p_val_table <- p_val_table %>% kable( booktabs = TRUE, digits=5, caption = "pvalue simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table$P5_t0_1<=0.05,'red','black'))
t_p_val_table_w <- p_val_table_w %>% kable( booktabs = TRUE, digits=5, caption = "pvalue Sonic abundance") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table_w$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table_w$P5_t0_1<=0.05,'red','black'))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t_p_val_table,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t_p_val_table_w,
    "\\end{minipage} 
\\end{table}"
))  


cont_table %>% kable( booktabs = TRUE, digits=3, caption = "simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "simple counts proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))

cont_table_w %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table_w %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cat('\\newpage')
```

```{r oncovar_4, echo=FALSE, results='asis'}
current_intsites <- intSitesTCGA_4
pandoc.header("OncoVar lvl 4", level = 1)

p_val_table <- get_pval_table(current_intsites)
cont_table <- get_cont_table(current_intsites)

p_val_table_w <- get_pval_table_w(current_intsites)
cont_table_w <- get_cont_table_w(current_intsites)


t_p_val_table <- p_val_table %>% kable( booktabs = TRUE, digits=5, caption = "pvalue simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table$P5_t0_1<=0.05,'red','black'))
t_p_val_table_w <- p_val_table_w %>% kable( booktabs = TRUE, digits=5, caption = "pvalue Sonic abundance") %>% kable_styling(latex_options = c("striped", "HOLD_position")) %>% column_spec(3,color=ifelse(p_val_table_w$P5_t0_2<=0.05,'red','black')) %>% column_spec(2,color=ifelse(p_val_table_w$P5_t0_1<=0.05,'red','black'))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t_p_val_table,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t_p_val_table_w,
    "\\end{minipage} 
\\end{table}"
))  


cont_table %>% kable( booktabs = TRUE, digits=3, caption = "simple counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "simple counts proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))

cont_table_w %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance counts") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table_w %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Sonic abundance proportions") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cat('\\newpage')
```


# Comments

* Using the Sonic Abundance counts better captures the differences between samples.
* The last timepoint of patient 5 has only 119 insertions, it might be too low for any meaningful conclusion.
* It is concerning that the two time zero samples are significantly different given most databases
* It would be hard to justify using a levels 2 or 3, either we use all genes that have some evidence of being oncogenes, or only the ones with the strongest evidence.

# To do

* I will merge the two time zero samples before collapsing replicates and use that as time zero
* UBR2 is a level 0 (no evidence for being oncogene) in OncoVar, I will test the effect of adding it to the oncogene lists

