---
title: "review_responses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library("tidyverse"))
library(kableExtra)
#source("fisher_exact_test.R")
#GTSP3310 -- p1_t72
#GTSP3311 -- p2_t72
#GTSP3438 -- p3_t72
#GTSP0886 -- p5_t0_1
#GTSP0890 -- p5_t0_2
#GTSP4185 -- P5_t70
```

# Reviewer 1

## fig3 question 1
**Add Gini, Shannon index and UC50**

We have added plots of the Gini and Shannon Index values for each subject to the supplementary material. (see sample table)

```{r ginifig, eval=FALSE, message=FALSE, include=FALSE}
source("gini.R")
```

## fig3 question 2

**The Authors should clearly declare the number of insertion sites retrieved for each of the analyzed samples.**

See sample table.



## fig3 question 4

**How to account for the known bias of lentiviral vectors to integrate in expressed genes**

We took the top 10% insertion sites by sonic abundance and constructed a contingency table to measure if oncogenes are over represented in the top 10% via Fisher exact test

### Patient 5 time 0, 1 

```{r load, include=FALSE}
source("fisher_exact_test.R")
driver_level_min=1
intSites <- readRDS("data/intSites_plus0.rds")
intSitesTCGA <- data.frame(readRDS(paste0('data/intSitesTCGA_',driver_level_min,'.rds')))
```


```{r fisher_Test_values, include=FALSE}
p5_t0_1_df <- get_sample_df('GTSP0886',intSitesTCGA)
p5_t0_2_df <- get_sample_df('GTSP0890',intSitesTCGA)
p5_t70_df <- get_sample_df('GTSP4185',intSitesTCGA)
p1_t72_df <- get_sample_df('GTSP3310',intSitesTCGA)
p2_t72_df <- get_sample_df('GTSP3311',intSitesTCGA)
p3_t72_df <- get_sample_df('GTSP3438',intSitesTCGA)
cont_table <-data.frame(
  "type" = c("onco_yes", "onco_no"),
  "P5_t0_1"=get_2sample_mat(p5_t0_1_df,p1_t72_df)$s1,
  "P5_t0_2"=get_2sample_mat(p5_t0_2_df,p1_t72_df)$s1,
  "P1_t72"=get_2sample_mat(p5_t0_1_df,p1_t72_df)$s2,
  "P2_t72"=get_2sample_mat(p5_t0_1_df,p2_t72_df)$s2,
  "P3_t72"=get_2sample_mat(p5_t0_1_df,p3_t72_df)$s2,
  "P5_t70"=get_2sample_mat(p5_t0_1_df,p5_t70_df)$s2
)
```

```{r pval_table, echo=FALSE}


p_val_table %>% kable( booktabs = TRUE, digits=3, caption = "Suplementary table XX") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
```

```{r cont_table1, echo=FALSE}
cont_table %>% kable( booktabs = TRUE, digits=3, caption = "Suplementary table XX") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
cont_table %>% mutate_if(is.numeric, ~ .x/sum(.)) %>% kable( booktabs = TRUE, digits=3, caption = "Suplementary table XX") %>% kable_styling(latex_options = c("striped", "HOLD_position"))
```


```{r fig_p5_t0_1, eval=FALSE}
sample_name <- "GTSP0886"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 5 time 0, 2 

```{r fig_p5_t0_2, eval=FALSE}
sample_name <- "GTSP0890"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 5 time 70 

```{r fig_p5_t70, eval=FALSE}
sample_name <- "GTSP4185"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 1 time 72 

```{r fig_p1_t72, eval=FALSE}
sample_name <- "GTSP3310"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 2 time 72 

```{r fig_p2_t72, eval=FALSE}
sample_name <- "GTSP3311"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 3 time 72 

```{r fig_p3_t72, eval=FALSE}
sample_name <- "GTSP3438"
sample_df <- get_sample_df(sample_name,intSites)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

## fig3 question 3

**Repeat the analysis using other gene classes and reviewed cancer database**

Using oncovar TCGA genes with driver level 1 or more. 3,487 total genes.

### Patient 5 time 0, 1

```{r fig_p5_t0_1_bis, eval=FALSE}
sample_name <- "GTSP0886"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 5 time 0, 2 

```{r fig_p5_t0_2_bis, eval=FALSE}
sample_name <- "GTSP0890"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 5 time 70 

```{r fig_p5_t70_bis, eval=FALSE}
sample_name <- "GTSP4185"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 1 time 72 

```{r fig_p1_t72_bis, eval=FALSE}
sample_name <- "GTSP3310"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 2 time 72 

```{r fig_p2_t72_bis, eval=FALSE}
sample_name <- "GTSP3311"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage

### Patient 3 time 72 

```{r fig_p3_t72_bis, eval=FALSE}
sample_name <- "GTSP3438"
sample_df <- get_sample_df(sample_name,intSitesTCGA)
sample_mat <- get_f_mat(sample_df)
fisher.test(sample_mat)
get_sample_figure(sample_df)
```
\newpage



## fig3 question 5
**Statistics for single genes are missing.**

We have published statistical methods based on scan statistics for identifying clusters of integration sites. Here we find xxx

## fig3 question 6
**Marking levels are between 0.01 and 0.08 for PBMCs. Then the \>1\% abundance. relative to the transformed cells is really 0.0008\% relative to all PBMCs. How does this affect the analysis of integration sites? The definition of dominant clone seems to be an overestimation.**

We have further developed the annotation of dominant clones in and added new analysis to the paper. We carried out a simulation xxx

## fig3 question 7
**Can we discriminate between relative abundance of 1\% and 2\%?**
This was already addressed in (Sherman et al. 2017)

